### 2015-04-14 src/paper1/paper1.tex
Fixed spelling mistakes

\small {\bf Abstract} Most evidence that the brain uses Bayesian inference to integrate noisy sensory signals optimally has been obtained by showing that the noise levels in each modality separately can predict performance in combined conditions. Such a forward approach is difficult to implement when the various signals cannot be measured in isolation, as in spatial orientation, which involves the processing of visual, somatosensory, and vestibular cues. Instead, we applied an inverse probabilistic approach, based on optimal observer theory. Our goal was to investigate whether the perceptual differences found when probing two different states ---body-in-space and head-in-space orientation--- can be reconciled by a shared scheme using all available sensory signals. Using a psychometric approach, seven human subjects were tested on two orientation estimates at tilts \textless 120\textdegree: perception of body tilt [subjective body tilt (SBT)] and perception of visual vertical [subjective visual vertical (SVV)]. In all subjects, the SBT was more accurate than the SVV, which showed substantial systematic errors for tilt angles beyond 60\textdegree. Variability increased with tilt angle in both tasks, but was consistently lower in the SVV. The sensory integration model fitted both datasets very nicely. A further experiment, in which supine subjects judged their head orientation relative to the body, independently confirmed the predicted head-on-body noise by the model. Model predictions based on the derived noise properties from the various modalities were also consistent with previously published deficits in vestibular and somatosensory patients. We conclude that Bayesian computations can account for the typical differences in spatial orientation judgments associated with different task requirements.

@@

Here, we used two psychophysical tasks---subjective body tilt (SBT) and subjective visual vertical (SVV) to quantify the two orientation estimates in a group of healthy subjects. Using an inverse probabilistic approach, we obtained stable solutions for the noise properties of the involved sensor systems. Independent measurements of neck noise confirmed the levels predicted by the model. Forward model predictions based on these noise properties were consistent with previously published deficits of bilateral vestibular and paraplegic patients, which would be difficult to explain otherwise. Our results suggest that Bayesian integration of multisensory information explains the major task-dependent features in spatial orientation perception.

@@

The SBT experiment served to obtain a psychometric measure of each subject's accuracy and precision of body-tilt perception at five body-tilt angles: upright (\siang{0}, \SBT{0} task) and \siang{45} and \siang{90} right side and left side down (\SBT{\textpm45} task and \SBT{\textpm90} task). Negative angles indicated left side down. We applied the method of constant stimuli, using a set of 10 equidistant body-tilt angles, centered on tentative estimates of the subject's \siang{0} (SBT0), \siang{45} (\SBT{45}), \siang{-45} (\SBT{-45}), \siang{90} (\SBT{90}), and \siang{-90} (SBT-90) body-tilt percept. The latter were determined in a few pilot trials that also served to familiarize the subject with the task, without providing a reference of the five respective orientations to be tested. Relative to the test angle, we used test angle intervals of \siang{3}, \siang{4}, and \siang{4} in the \SBT{0}, \SBT{\textpm45}, and \SBT{\textpm90} tasks, respectively. Body-tilt angles were tested 14 times in random order, yielding 140 responses for each psychometric curve.

To perform the psychophysical SBT experiments, two methodological problems had to be solved. The first relates to the number of experimental sessions that we could reasonably ask subjects to perform. We realized that returning the subject to upright for reorientation after each trial would require too large a number of experimental sessions. Our overriding concern was that starting each trial from upright would confound the \SBT{0} task in the sense that subjects could then simply notice the change in chair position. To prevent this, we always inserted a detour rotation before moving the subject to the test angle in a given trial. The detour, always to a tilt position clearly outside the psychometric test range, served to reset the subject's memory of the previous tilt position. These detour angles were chosen randomly from a range at \siang{30-40} clockwise (CW) and counterclockwise (CCW) from the presumed threshold. As an illustration, \figref{p1:fig2} shows how the subject was moved from one trial to the next in the course of an \SBT{90} experiment. Detour angles preceding each test angle were taken from the CW and the CCW detour range in equal proportions. An analysis of trial history effects indicated that detour angles did not affect the judgment in the subsequent trial ($p > 0.05$).

@@

    \caption{Tilt paradigm in \SBT{90} task. T1, T2, T3, Test angles at which the subject was prompted with a beep signal (*) to indicate whether body orientation was CW or CCW from the instructed reference orientation (i.e., \siang{90} in this example). D1, D2, Detour angles randomly drawn from detour range (\siang{30-40} CW and CCW from center of test range). Rotations from detour (D) to test (T) angle were performed in a noisy fashion (see \nameref{p1:sec:methods}, \nameref{p1:sec:methods_sbt}).}
    \label{p1:fig2}
\end{wrapfigure}

@@

The same subjects were also tested in a series of SVV experiments. Part of this dataset (four subjects) has been published previously as part of a larger dataset on visual verticality perception \cite{devrijer2009}. Data in the other three subjects were collected anew. Here we provide a brief summary of the paradigm. SVV was tested at nine roll-tilt angles, ranging from -120 to \siang{120} at \siang{30} intervals. A luminous line (angular subtend, \siang{20}), polarized with a bright dot at one end, was mounted in front of the subject. The line's rotation axis coincided with the chair rotation axis. In each experimental run, the subject was rotated from upright to the chosen test angle at a constant angular velocity of \siang[/s]{30}. After a 30 \si{\second} waiting period that allowed canal effects to subside, a luminous line was briefly flashed (20 \si{\milli\second}), and the subject indicated whether its orientation in space was CW or CCW from the perceived direction of gravity. The line orientation was selected randomly from a set of 11 line orientations. After all line orientations had been tested, the subject was rotated back to upright, and room lights were turned on. Positive and negative body-tilt angles were alternated regularly. As in the SBT experiment, we used the method of constant stimuli. The set of 11 line orientations was centered on a coarse estimate of the SVV threshold at each tilt angle. We used orientation intervals of \siang{3}, except for upright, where intervals of \siang{2} were taken. Each set of line orientations was tested in random order in 12 experimental runs, thus yielding a total of 132 responses for each psychometric curve. SVV data were collected in a total of five 45 min sessions per subject.

@@

To obtain an estimate of the orientation of the body in space, the brain can use ``direct'' sensory information from body sensors ($\hat{B}_S$), such as tactile receptors in the skin or so-called graviceptors in the trunk \cite{mittelstaedt1997, mittelstaedt1998, vaitl2002}. Alternatively, an ``indirect'' pathway, involving a reference frame transformation, can also provide a body-in-space estimate. For this purpose, sensory head-in-space information, provided by the otoliths, must be combined with information about head-on-body orientation, provided by proprioceptive signals from the neck ($\hat{B}_{SI} = \hat{H}_S - \hat{H}_B$). Because the sensors are contaminated by noise, the direct and indirect signals can be represented as Gaussian probability distributions with mean values of $\hat{B}_S$ and $\hat{H}_S - \hat{H}_B$, and variance levels of $\sigma_{BS}^2$ and $\sigma_{HS}^2 + \sigma_{HB}^2$, respectively. Theoretically, as shown in \figref{p1:fig1}, the brain could also use prior information about body-in-space orientation in the computation of the body-in-space estimate. The effect of including a prior on the SBT (centered on upright) would be a systematic error of underestimation at larger tilt angles. However, neither previous findings \cite{mittelstaedt1983, mast1996, jarchow1999, vanbeuzekom2001} nor the present results (\figref[ and ]{p1:fig3,p1:fig4}), showed such systematic errors across subjects. In modeling terms, this indicates a uniform (uninformative) prior, which corresponds to a weight of 0. Accordingly, a statistically optimal estimate of body-in-space orientation ($\tilde{B}_S$) is then given by the peak of the Gaussian distribution that results from the multiplication of the two distributions representing the direct and indirect sensory pathways. It follows that

@@

The scheme applies a similar sensory signal processing strategy to estimate the orientation of head in space, $\tilde{H}_S$, used in the SVV. A direct estimate of head-in-space orientation is provided by the head-in-space sensors ($\hat{H}_S$), and an indirect estimate is obtained by a reference frame transformation of the body-in-space signal ($\hat{B}_S$) by adding the head-on-body estimate ($\hat{H}_B$), provided by the neck sensors ($\hat{H}_{SI} = \hat{B}_S + \hat{H}_B$). Again, direct and indirect pathway signals are represented by two Gaussian probability distributions, with mean values of $\hat{H}_S$ and $\hat{B}_S + \hat{H}_B$, respectively, and corresponding variances of $\sigma^2_{HS}$ and $\sigma^2_{BS} + \sigma^2_{HB}$. In the computation of the head-in-space estimate, to account for systematic errors \cite{macneilage2007, devrijer2008}, it is further assumed that the brain uses prior knowledge about head-in-space orientation, which entails that small head-tilt angles are considered more probable than large tilts. Mathematically, the prior is represented by a Gaussian distribution that is centered at \siang{0} head tilt ($H_{SP} = 0^\circ$) with a variance of $\sigma^2_{HSP}$. Note that, in our scheme, the head-in-space prior, which contributes to the SVV computations, does not affect the body-in-space estimate. Integration of the direct and indirect sensory pathways and prior knowledge is performed by multiplication of the three Gaussian distributions. The peak of the resulting posterior distribution represents the optimal estimate of head-in-space orientation ($\tilde{H}_S$), which is given by the following:

@@

In this equation, $w_{HD}$, $w_{HI}$, and $w_{HP}$ (which add up to one) represent the weights of the direct and indirect pathways and the prior, respectively, which are proportional to the relative precision of the sensory signals and the width of the prior. \eqnref{p1:eqn6} would result in an accurate estimate of $\tilde{H}_S$, if all three pathways were accurate by themselves. However, because the prior is centered on zero ($H_{SP} = 0\degree$), it introduces more and more bias toward upright, as head tilt increases further. Thus, optimization in terms of variance has a downside by causing underestimation of the actual head tilt. The amount of underestimation depends on the width of the prior and the reliability of the sensory inputs.

@@

In addition to these ``parameters of interest'', the data were preprocessed before model fitting by applying mean correction \cite{mcguire2009}. Mean correction was performed to remove systematic errors in the SBT and the asymmetries in the SVV between CW and CCW tilt angles. Because the model is inherently left-right symmetric, it would try to account for differences in SVV bias between equal but opposite tilt angles by falsely increasing the variance. Likewise, because the model assumes that there is no bias in the SBT, it would try to explain any slight deviation from zero by excessively increasing the variance. The asymmetry in the SVV, if any, and a nonzero SBT bias, if any, are captured by fixed parameters of noninterest (n = 9) in the model fits. Thus, for the SBT data, one bias correction term was needed for each tilt angle (yielding five parameters of noninterest), and for the SVV data, one correction parameter was needed for each pair of equal but opposite tilt angles (yielding four parameters of noninterest). We emphasize that the nine parameters of noninterest are not free-fit parameters because they are not optimized by the model. So, although technically our number of free parameters amounts to a total of 16, only seven were determined by fitting the model.

In total, the seven free parameters of the model had to account for 149 data points, spread across various tilt angles, with each data point reflecting a proportion of CW responses based on either 14 (for the SBT) or 11 (for the SVV) experimental forced-choice CW/CCW responses. We fitted the model by maximizing the likelihood of the data [maximum likelihood estimation (MLE)], in relation to the set of six model parameters ($a_{HS}$, $b_{HS}$, $\sigma_{HSP}$, $\sigma_{BS}$, $\sigma_{HB}$, and $A_{OCR}$) and lapse rate ($\lambda$). Optimal parameter values were obtained by minimizing the negative likelihood function using the Matlab function ``fmincon'' \cite{devrijer2008, mcguire2009}. Simulations confirmed that the inverse modeling approach was not sensitive to overfitting. SDs of the best-fit parameters were obtained by performing 1999 bootstrap runs. For each run, we constructed 149 data points (reflecting the size of the dataset), each of which was obtained by random sampling with replacement from the original dataset. The model was fit to this new dataset. The distribution of model parameters across all runs was used to derive the 68.2\% confidence interval of each parameter.

We emphasize that the model fit provided an estimate of the proprioceptive variance of the neck ($\sigma_{HB}$), even though the head-on-body signal was not directly manipulated during the experiment. Nevertheless, this signal, as sensed by the neck proprioceptors, is essential to implement the reference frame transformation from the body-in-space to the head-in-space signal, and vice versa. Because the neck signal is noisy, these reference frame transformations induce neck-related noise in the original $B_S$ and $H_S$ signals, even when the head and body are aligned. Because the SBT and SVV tasks require different reference frame transformations, they depend differently on the noise properties of the three sensory systems (body and neck sensors and otoliths). By solving the inverse problem, the noise properties of the three sensory systems, as well as the other fit parameters, can be determined. Finally, we note that the inverse problem can only be solved using both tasks at multiple tilt angles; just using a single task (SVV or SBT, not both) would have made this problem intractable.

@@

Furthermore, we evaluated our mechanistic model in comparison with a pure descriptive model of the same dataset based on separate psychometric accounts, each with three free parameters, at the five SBT and nine SVV angles. We used the Bayesian information criterion (BIC) for model comparison. BIC provides a measure of the adequacy of the model fit and corrects for the number of parameters. The BIC is defined as $BIC = -2 \log(L) + k \cdot \log(n)$, in which $L$ is the total likelihood of the data given the model, $k$ the number of free parameters, and $n$ the number of data points to be explained. The number of free parameters is 42 [14 psychometric curves $\times$ 3 parameters ($\mu$, $\sigma$, and $\lambda$)] for the psychometric curves, whereas for the mechanistic Bayesian model, the number of free parameters is seven. A more appropriate model is characterized by a lower BIC value.

@@

In complete darkness, subjects were first rotated at a constant angular velocity of \siang[/s]{$\leq15$} to a random detour angle similar to the idea shown in \figref{p1:fig2}. The head then moved to a randomly chosen position within the test range with a very slow speed (\siang[/s]{$<0.2$}) such that the test angle was reached within 20 \si{\second}. Meanwhile, auditory white noise was presented to the subjects through earphones to mask any auditory cues generated by the moving platform. After arrival at the test angle, the auditory noise was interrupted, signaling the subject to indicate whether head-on-body orientation was CW or CCW relative to the body midline, using a toggle switch. The subject was then rotated at a constant velocity to a new randomly chosen detour angle, and the above procedure was repeated. Each test angle was repeated 10 times, yielding a total number of 110 responses in each subject. Psychometric data were quantified by fitting a cumulative Gaussian function (see above, \eqnref{p1:eqn1}). The width of the curve, $\sigma^2$, inversely related to precision, serves as an independent measure of the subject's variability of the head-on-body estimate and was compared with the model prediction.

@@

The SBT experiment, performed in seven subjects, tested the accuracy and precision of SBT percepts, near upright (SBT0), at \siang{45} and \siang{90} right side down (\SBT{45} and \SBT{90}), and at \siang{45} and \siang{90} left side down (\SBT{-45} and \SBT{-90}). The same subjects also performed the SVV experiment, tested at nine roll-tilt angles, ranging from \siang{-120} to \siang{120} in \siang{30} intervals. \figref{p1:fig3} shows the results of a typical subject (S1) in both tasks. The top panels show the proportion of CW responses for the five SBT tasks, relative to the reference orientation. For an ideal observer, all psychometric functions would resemble a step centered at zero. Across the five reference orientations (\siang{0}, \siang{{\textpm}45}, or \siang{{\textpm}90}), the psychometric data indicate underestimations and overestimations of perceived body angle, but no consistent bias, which resembles previous reports \cite{mittelstaedt1983, mast1996, jarchow1999, vanbeuzekom2001} that body-tilt perception is accurate on average. We fitted psychometric curves through these data (see \nameref{p1:sec:methods}, Eq. \ref{p1:eqn1}), to obtain estimates for the mean ($\mu$), SD ($\sigma$), and lapse rate ($\lambda$). Parameter $\mu$ is a measure for the accuracy of the subject's body-tilt percept. Perceptual variability, inversely related to precision, is reflected by $\sigma^2$, whereas the lapse rate ($\lambda = 0.06$) accounts for stimulus-independent errors \cite{wichmann2001}. In all five SBT tasks, the $\mu$ values are relatively close to the veridical reference orientation (0\textdegree, 45\textdegree, or 90\textdegree), i.e., errors are \textless5\textdegree. The psychometric fits further show that variability is lower in the \SBT{0} task, with $\sigma \approx 4\degree$, than in the \SBT{\textpm45} and \SBT{\textpm90} task, where $\sigma \approx 10\degree$.

@@

The two bottom rows of \figref{p1:fig4} summarize our SVV data across the entire tilt range. Accuracy is close to perfect at upright orientation in all subjects, with mean values ranging between \siang{0.1} and 2.8\textdegree. For tilts $\ge60\degree$, all subjects show systematic SVV errors (biases) of undercompensation, ranging up to maximum values close to 60\textdegree. Three subjects (S1, S2, and S3) also show slight errors of overcompensation in the smallest tilt range (\textless60\textdegree). The variability in the SVV is \siang{\textless3.0} for upright, which is consistently lower than in the tilted conditions, where variability reaches values ranging up to 8\textdegree.

@@

Instead of an overall dominance of body receptors in the direct pathway, the model implies that it is actually the indirect pathway, carrying the otolith signal, that dominates the behaviorally important range near upright. In most of our subjects (S1, S3, S5, S6, and S7), it was only when the otoliths became less reliable, at larger tilts, that the body sensors (direct pathway) got the upper hand ($w_{BD} > 0.5$).

@@

In this study, we made intrasubject comparisons of the accuracy (bias) and precision (inverse variance) characteristics in two spatial orientation tasks: SBT and SVV. The main experimental findings were as follows: (1) the SBT is more accurate than the SVV, (2) the SBT is less precise than the SVV, and (3) both SBT and SVV precision are smaller in tilted conditions than near upright. Under the assumption of optimality, a Bayesian model of sensory integration could account very well for these findings. Independent measurements, in supine subjects, of head-on-body variance confirmed the predicted value.

@@

The architecture of the model, as far as the reference frame transformations and the sensory integration is concerned, follows entirely from the principles of Bayesian inference. However, to account for our major findings and intersubject differences, we made two less straightforward assumptions. First, to explain the increased variability in both tasks at \siang{90} tilt, we allowed for the possibility that the otoliths become more noisy with increases in tilt. Second, we hypothesized that prior knowledge is used in the visual vertical but not in body-tilt perception. Can these assumptions be justified on physiological and rational grounds?

>>>>

\small {\bf Abstract} Most evidence that the brain uses Bayesian inference to integrate noisy sensory signals optimally has been obtained by showing that the noise levels in each modality separately can predict performance in combined conditions. Such a forward approach is difficult to implement when the various signals cannot be measured in isolation, as in spatial orientation, which involves the processing of visual, somatosensory, and vestibular cues. Instead, we applied an inverse probabilistic approach, based on optimal observer theory. Our goal was to investigate whether the perceptual differences found when probing two different states -- body-in-space and head-in-space orientation -- can be reconciled by a shared scheme using all available sensory signals. Using a psychometric approach, seven human subjects were tested on two orientation estimates at tilts \textless 120\textdegree: perception of body tilt [subjective body tilt (SBT)] and perception of visual vertical [subjective visual vertical (SVV)]. In all subjects, the SBT was more accurate than the SVV, which showed substantial systematic errors for tilt angles beyond 60\textdegree. Variability increased with tilt angle in both tasks, but was consistently lower in the SVV. The sensory integration model fitted both datasets very nicely. A further experiment, in which supine subjects judged their head orientation relative to the body, independently confirmed the predicted head-on-body noise by the model. Model predictions based on the derived noise properties from the various modalities were also consistent with previously published deficits in vestibular and somatosensory patients. We conclude that Bayesian computations can account for the typical differences in spatial orientation judgments associated with different task requirements.

@@

Here, we used two psychophysical tasks -- subjective body tilt (SBT) and subjective visual vertical (SVV) to quantify the two orientation estimates in a group of healthy subjects. Using an inverse probabilistic approach, we obtained stable solutions for the noise properties of the involved sensor systems. Independent measurements of neck noise confirmed the levels predicted by the model. Forward model predictions based on these noise properties were consistent with previously published deficits of bilateral vestibular and paraplegic patients, which would be difficult to explain otherwise. Our results suggest that Bayesian integration of multisensory information explains the major task-dependent features in spatial orientation perception.

@@

The SBT experiment served to obtain a psychometric measure of each subject's accuracy and precision of body-tilt perception at five body-tilt angles: upright (\siang{0}, \SBT{0} task) and \siang{45} and \siang{90} right side and left side down (\SBT{\textpm45} task and \SBT{\textpm90} task). Negative angles indicated left side down. We applied the method of constant stimuli, using a set of 10 equidistant body-tilt angles, centred on tentative estimates of the subject's \siang{0} (SBT0), \siang{45} (\SBT{45}), \siang{-45} (\SBT{-45}), \siang{90} (\SBT{90}), and \siang{-90} (SBT-90) body-tilt percept. The latter were determined in a few pilot trials that also served to familiarise the subject with the task, without providing a reference of the five respective orientations to be tested. Relative to the test angle, we used test angle intervals of \siang{3}, \siang{4}, and \siang{4} in the \SBT{0}, \SBT{\textpm45}, and \SBT{\textpm90} tasks, respectively. Body-tilt angles were tested 14 times in random order, yielding 140 responses for each psychometric curve.

To perform the psychophysical SBT experiments, two methodological problems had to be solved. The first relates to the number of experimental sessions that we could reasonably ask subjects to perform. We realised that returning the subject to upright for reorientation after each trial would require too large a number of experimental sessions. Our overriding concern was that starting each trial from upright would confound the \SBT{0} task in the sense that subjects could then simply notice the change in chair position. To prevent this, we always inserted a detour rotation before moving the subject to the test angle in a given trial. The detour, always to a tilt position clearly outside the psychometric test range, served to reset the subject's memory of the previous tilt position. These detour angles were chosen randomly from a range at \siang{30-40} clockwise (CW) and counterclockwise (CCW) from the presumed threshold. As an illustration, \figref{p1:fig2} shows how the subject was moved from one trial to the next in the course of an \SBT{90} experiment. Detour angles preceding each test angle were taken from the CW and the CCW detour range in equal proportions. An analysis of trial history effects indicated that detour angles did not affect the judgment in the subsequent trial ($p > 0.05$).

@@

    \caption{Tilt paradigm in \SBT{90} task. T1, T2, T3, Test angles at which the subject was prompted with a beep signal (*) to indicate whether body orientation was CW or CCW from the instructed reference orientation (i.e., \siang{90} in this example). D1, D2, Detour angles randomly drawn from detour range (\siang{30-40} CW and CCW from centre of test range). Rotations from detour (D) to test (T) angle were performed in a noisy fashion (see \nameref{p1:sec:methods}, \nameref{p1:sec:methods_sbt}).}
    \label{p1:fig2}
\end{wrapfigure}

@@

The same subjects were also tested in a series of SVV experiments. Part of this dataset (four subjects) has been published previously as part of a larger dataset on visual verticality perception \cite{devrijer2009}. Data in the other three subjects were collected anew. Here we provide a brief summary of the paradigm. SVV was tested at nine roll-tilt angles, ranging from -120 to \siang{120} at \siang{30} intervals. A luminous line (angular subtend, \siang{20}), polarised with a bright dot at one end, was mounted in front of the subject. The line's rotation axis coincided with the chair rotation axis. In each experimental run, the subject was rotated from upright to the chosen test angle at a constant angular velocity of \siang[/s]{30}. After a 30 \si{\second} waiting period that allowed canal effects to subside, a luminous line was briefly flashed (20 \si{\milli\second}), and the subject indicated whether its orientation in space was CW or CCW from the perceived direction of gravity. The line orientation was selected randomly from a set of 11 line orientations. After all line orientations had been tested, the subject was rotated back to upright, and room lights were turned on. Positive and negative body-tilt angles were alternated regularly. As in the SBT experiment, we used the method of constant stimuli. The set of 11 line orientations was centred on a coarse estimate of the SVV threshold at each tilt angle. We used orientation intervals of \siang{3}, except for upright, where intervals of \siang{2} were taken. Each set of line orientations was tested in random order in 12 experimental runs, thus yielding a total of 132 responses for each psychometric curve. SVV data were collected in a total of five 45 min sessions per subject.

@@

To obtain an estimate of the orientation of the body in space, the brain can use ``direct'' sensory information from body sensors ($\hat{B}_S$), such as tactile receptors in the skin or so-called graviceptors in the trunk \cite{mittelstaedt1997, mittelstaedt1998, vaitl2002}. Alternatively, an ``indirect'' pathway, involving a reference frame transformation, can also provide a body-in-space estimate. For this purpose, sensory head-in-space information, provided by the otoliths, must be combined with information about head-on-body orientation, provided by proprioceptive signals from the neck ($\hat{B}_{SI} = \hat{H}_S - \hat{H}_B$). Because the sensors are contaminated by noise, the direct and indirect signals can be represented as Gaussian probability distributions with mean values of $\hat{B}_S$ and $\hat{H}_S - \hat{H}_B$, and variance levels of $\sigma_{BS}^2$ and $\sigma_{HS}^2 + \sigma_{HB}^2$, respectively. Theoretically, as shown in \figref{p1:fig1}, the brain could also use prior information about body-in-space orientation in the computation of the body-in-space estimate. The effect of including a prior on the SBT (centred on upright) would be a systematic error of underestimation at larger tilt angles. However, neither previous findings \cite{mittelstaedt1983, mast1996, jarchow1999, vanbeuzekom2001} nor the present results (\figref[ and ]{p1:fig3,p1:fig4}), showed such systematic errors across subjects. In modelling terms, this indicates a uniform (uninformative) prior, which corresponds to a weight of 0. Accordingly, a statistically optimal estimate of body-in-space orientation ($\tilde{B}_S$) is then given by the peak of the Gaussian distribution that results from the multiplication of the two distributions representing the direct and indirect sensory pathways. It follows that

@@

The scheme applies a similar sensory signal processing strategy to estimate the orientation of head in space, $\tilde{H}_S$, used in the SVV. A direct estimate of head-in-space orientation is provided by the head-in-space sensors ($\hat{H}_S$), and an indirect estimate is obtained by a reference frame transformation of the body-in-space signal ($\hat{B}_S$) by adding the head-on-body estimate ($\hat{H}_B$), provided by the neck sensors ($\hat{H}_{SI} = \hat{B}_S + \hat{H}_B$). Again, direct and indirect pathway signals are represented by two Gaussian probability distributions, with mean values of $\hat{H}_S$ and $\hat{B}_S + \hat{H}_B$, respectively, and corresponding variances of $\sigma^2_{HS}$ and $\sigma^2_{BS} + \sigma^2_{HB}$. In the computation of the head-in-space estimate, to account for systematic errors \cite{macneilage2007, devrijer2008}, it is further assumed that the brain uses prior knowledge about head-in-space orientation, which entails that small head-tilt angles are considered more probable than large tilts. Mathematically, the prior is represented by a Gaussian distribution that is centred at \siang{0} head tilt ($H_{SP} = 0^\circ$) with a variance of $\sigma^2_{HSP}$. Note that, in our scheme, the head-in-space prior, which contributes to the SVV computations, does not affect the body-in-space estimate. Integration of the direct and indirect sensory pathways and prior knowledge is performed by multiplication of the three Gaussian distributions. The peak of the resulting posterior distribution represents the optimal estimate of head-in-space orientation ($\tilde{H}_S$), which is given by the following:

@@

In this equation, $w_{HD}$, $w_{HI}$, and $w_{HP}$ (which add up to one) represent the weights of the direct and indirect pathways and the prior, respectively, which are proportional to the relative precision of the sensory signals and the width of the prior. \eqnref{p1:eqn6} would result in an accurate estimate of $\tilde{H}_S$, if all three pathways were accurate by themselves. However, because the prior is centred on zero ($H_{SP} = 0\degree$), it introduces more and more bias toward upright, as head tilt increases further. Thus, optimization in terms of variance has a downside by causing underestimation of the actual head tilt. The amount of underestimation depends on the width of the prior and the reliability of the sensory inputs.

@@

In addition to these ``parameters of interest'', the data were preprocessed before model fitting by applying mean correction \cite{mcguire2009}. Mean correction was performed to remove systematic errors in the SBT and the asymmetries in the SVV between CW and CCW tilt angles. Because the model is inherently left-right symmetric, it would try to account for differences in SVV bias between equal but opposite tilt angles by falsely increasing the variance. Likewise, because the model assumes that there is no bias in the SBT, it would try to explain any slight deviation from zero by excessively increasing the variance. The asymmetry in the SVV, if any, and a nonzero SBT bias, if any, are captured by fixed parameters of non-interest (n = 9) in the model fits. Thus, for the SBT data, one bias correction term was needed for each tilt angle (yielding five parameters of non-interest), and for the SVV data, one correction parameter was needed for each pair of equal but opposite tilt angles (yielding four parameters of non-interest). We emphasise that the nine parameters of non-interest are not free-fit parameters because they are not optimised by the model. So, although technically our number of free parameters amounts to a total of 16, only seven were determined by fitting the model.

In total, the seven free parameters of the model had to account for 149 data points, spread across various tilt angles, with each data point reflecting a proportion of CW responses based on either 14 (for the SBT) or 11 (for the SVV) experimental forced-choice CW/CCW responses. We fitted the model by maximizing the likelihood of the data [maximum likelihood estimation (MLE)], in relation to the set of six model parameters ($a_{HS}$, $b_{HS}$, $\sigma_{HSP}$, $\sigma_{BS}$, $\sigma_{HB}$, and $A_{OCR}$) and lapse rate ($\lambda$). Optimal parameter values were obtained by minimizing the negative likelihood function using the Matlab function ``fmincon'' \cite{devrijer2008, mcguire2009}. Simulations confirmed that the inverse modelling approach was not sensitive to overfitting. SDs of the best-fit parameters were obtained by performing 1999 bootstrap runs. For each run, we constructed 149 data points (reflecting the size of the dataset), each of which was obtained by random sampling with replacement from the original dataset. The model was fit to this new dataset. The distribution of model parameters across all runs was used to derive the 68.2\% confidence interval of each parameter.

We emphasise that the model fit provided an estimate of the proprioceptive variance of the neck ($\sigma_{HB}$), even though the head-on-body signal was not directly manipulated during the experiment. Nevertheless, this signal, as sensed by the neck proprioceptors, is essential to implement the reference frame transformation from the body-in-space to the head-in-space signal, and vice versa. Because the neck signal is noisy, these reference frame transformations induce neck-related noise in the original $B_S$ and $H_S$ signals, even when the head and body are aligned. Because the SBT and SVV tasks require different reference frame transformations, they depend differently on the noise properties of the three sensory systems (body and neck sensors and otoliths). By solving the inverse problem, the noise properties of the three sensory systems, as well as the other fit parameters, can be determined. Finally, we note that the inverse problem can only be solved using both tasks at multiple tilt angles; just using a single task (SVV or SBT, not both) would have made this problem intractable.

@@

Furthermore, we evaluated our mechanistic model in comparison with a pure descriptive model of the same dataset based on separate psychometric accounts, each with three free parameters, at the five SBT and nine SVV angles. We used the Bayesian information criterion (BIC) for model comparison. BIC provides a measure of the adequacy of the model fit and corrects for the number of parameters. The BIC is defined as $BIC = -2 \log(L) + k \cdot \log(n)$, in which $L$ is the total likelihood of the data given the model, $k$ the number of free parameters, and $n$ the number of data points to be explained. The number of free parameters is 42 [14 psychometric curves $\times$ 3 parameters ($\mu$, $\sigma$, and $\lambda$)] for the psychometric curves, whereas for the mechanistic Bayesian model, the number of free parameters is seven. A more appropriate model is characterised by a lower BIC value.

@@

In complete darkness, subjects were first rotated at a constant angular velocity of \siang[/s]{$\leq15$} to a random detour angle similar to the idea shown in \figref{p1:fig2}. The head then moved to a randomly chosen position within the test range with a very slow speed (\siang[/s]{$<0.2$}) such that the test angle was reached within 20 \si{\second}. Meanwhile, auditory white noise was presented to the subjects through earphones to mask any auditory cues generated by the moving platform. After arrival at the test angle, the auditory noise was interrupted, signalling the subject to indicate whether head-on-body orientation was CW or CCW relative to the body midline, using a toggle switch. The subject was then rotated at a constant velocity to a new randomly chosen detour angle, and the above procedure was repeated. Each test angle was repeated 10 times, yielding a total number of 110 responses in each subject. Psychometric data were quantified by fitting a cumulative Gaussian function (see above, \eqnref{p1:eqn1}). The width of the curve, $\sigma^2$, inversely related to precision, serves as an independent measure of the subject's variability of the head-on-body estimate and was compared with the model prediction.

@@

The SBT experiment, performed in seven subjects, tested the accuracy and precision of SBT percepts, near upright (SBT0), at \siang{45} and \siang{90} right side down (\SBT{45} and \SBT{90}), and at \siang{45} and \siang{90} left side down (\SBT{-45} and \SBT{-90}). The same subjects also performed the SVV experiment, tested at nine roll-tilt angles, ranging from \siang{-120} to \siang{120} in \siang{30} intervals. \figref{p1:fig3} shows the results of a typical subject (S1) in both tasks. The top panels show the proportion of CW responses for the five SBT tasks, relative to the reference orientation. For an ideal observer, all psychometric functions would resemble a step centred at zero. Across the five reference orientations (\siang{0}, \siang{{\textpm}45}, or \siang{{\textpm}90}), the psychometric data indicate underestimations and overestimations of perceived body angle, but no consistent bias, which resembles previous reports \cite{mittelstaedt1983, mast1996, jarchow1999, vanbeuzekom2001} that body-tilt perception is accurate on average. We fitted psychometric curves through these data (see \nameref{p1:sec:methods}, Eq. \ref{p1:eqn1}), to obtain estimates for the mean ($\mu$), SD ($\sigma$), and lapse rate ($\lambda$). Parameter $\mu$ is a measure for the accuracy of the subject's body-tilt percept. Perceptual variability, inversely related to precision, is reflected by $\sigma^2$, whereas the lapse rate ($\lambda = 0.06$) accounts for stimulus-independent errors \cite{wichmann2001}. In all five SBT tasks, the $\mu$ values are relatively close to the veridical reference orientation (0\textdegree, 45\textdegree, or 90\textdegree), i.e., errors are \textless5\textdegree. The psychometric fits further show that variability is lower in the \SBT{0} task, with $\sigma \approx 4\degree$, than in the \SBT{\textpm45} and \SBT{\textpm90} task, where $\sigma \approx 10\degree$.

@@

The two bottom rows of \figref{p1:fig4} summarise our SVV data across the entire tilt range. Accuracy is close to perfect at upright orientation in all subjects, with mean values ranging between \siang{0.1} and 2.8\textdegree. For tilts $\ge60\degree$, all subjects show systematic SVV errors (biases) of undercompensation, ranging up to maximum values close to 60\textdegree. Three subjects (S1, S2, and S3) also show slight errors of overcompensation in the smallest tilt range (\textless60\textdegree). The variability in the SVV is \siang{\textless3.0} for upright, which is consistently lower than in the tilted conditions, where variability reaches values ranging up to 8\textdegree.

@@

Instead of an overall dominance of body receptors in the direct pathway, the model implies that it is actually the indirect pathway, carrying the otolith signal, that dominates the behaviourally important range near upright. In most of our subjects (S1, S3, S5, S6, and S7), it was only when the otoliths became less reliable, at larger tilts, that the body sensors (direct pathway) got the upper hand ($w_{BD} > 0.5$).

@@

In this study, we made intra-subject comparisons of the accuracy (bias) and precision (inverse variance) characteristics in two spatial orientation tasks: SBT and SVV. The main experimental findings were as follows: (1) the SBT is more accurate than the SVV, (2) the SBT is less precise than the SVV, and (3) both SBT and SVV precision are smaller in tilted conditions than near upright. Under the assumption of optimality, a Bayesian model of sensory integration could account very well for these findings. Independent measurements, in supine subjects, of head-on-body variance confirmed the predicted value.

@@

The architecture of the model, as far as the reference frame transformations and the sensory integration is concerned, follows entirely from the principles of Bayesian inference. However, to account for our major findings and inter-subject differences, we made two less straightforward assumptions. First, to explain the increased variability in both tasks at \siang{90} tilt, we allowed for the possibility that the otoliths become more noisy with increases in tilt. Second, we hypothesised that prior knowledge is used in the visual vertical but not in body-tilt perception. Can these assumptions be justified on physiological and rational grounds?
